Climatic warming strengthens a positive feedback between alpine shrubs and fire (Supplementary Material)

### 2016-03-08 manuscript/supplementary.Rnw
Put \label under the caption
More edits to supp mat

\section*{Study site.}
Research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals\cite{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003\cite{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests\cite{Williams:2008ba}.

\section*{Open Top Chamber Experiment.}
Open Top Chamber Experiment
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.

In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}. Here, we used fire to create bare ground because 1) Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and 2) in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Flannigan:2009kl, Westerling:2011jp}. Thus, to forecast likely changes in shrub abundance we need to know how warming affects growth and mortality rates of shrub seedlings growing in burnt sites.

We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient seedling numbers from species occurring in both open and closed heathland. We focused on \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) and \textit{Phebalium squamulosum} (Rutaceae; a tall shrub because they are all: obligate seeders post-fire (i.e. fire killed), common in the Australian Alps, and under global warming, likely candidates to invade non-shrubby plant communities such as alpine grasslands and herbfields. For each species, we collected a range of seedling heights (2 – 12 cm) using a forestry planter (12.5 cm deep by 5.5 cm wide). The size of the planter was selected to minimize root damage across the range of seedling sizes and to allow for the extraction of a constant volume (378 cm$^3$) of intact soil and roots. The planter also minimized soil disturbance during seedling extraction and replanting.

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea seedlings} were transplanted into various sized inter-tussock gaps surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. S8). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

To simulate near-term warmer conditions predicted under IPCC\cite{IPCC:2013tu}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}. OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm. The chambers were constructed following the International Tundra Experiment (ITEX) protocols\cite{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (< 5\%; Solar Components Corporation, USA).

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) randomly installed into four control plots and four OTC plots. Each Micro Station measured ambient temperature, relative humidity (both at 10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013tu} (Supplementary Fig. S5-7). OTCs passively raised mean daily ambient and soil temperatures by approximately 1 \textdegree C and 0.9 \textdegree C ($\pm$ 0.05 95\% CI) and marginally lowered both soil moisture (- 0.21\% $\pm$ 0.11) and relative humidity (- 1.2\% $\pm$ 0.17). OTCs also increased minimum daily ambient and soil temperatures by 0.86 \textdegree C ($\pm$ 0.06 95\% CI) and 1.6 ($\pm$ 0.12 95\% CI) respectively. Maximum temperatures were also strongly influenced by OTCs with ambient and soil maximum temperatures increasing by 2.4 \textdegree C ($\pm$  0.16 95\% CI) and 0.95 \textdegree C ($\pm$  0.42 95\% CI), respectively. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\section*{Seedling survey.}
We measured seedling density and maximum heights in open heathland across the Bogong High Plains. The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (a measure of plant available water), adult density, fire, and, if burnt, fire severity, influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland. A secondary aim was to validate growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.

The study used 40 open heathland sites\cite{Camac:2013eb} established immediately post-2003 fires \cite{Williams:2006tl}. All sites have not been grazed by cattle since the 2003 fires. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years\cite{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitudinal from 1660 m to 1860 m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2012-13, 9-years post-fire, seedling density in each site was estimated using 40 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots \texttimes 4 transects) per site, subsampling an area of 2000 m$^2$.

Different variables were recorded at the three sampling levels (site, transect and plot). For each site, altitude and TWI were derived from a 30 m resolution digital elevation model of the Bogong High Plains. TWI, which approximates the relative plant-available water based on a site’s position in the landscape\cite{Moore:1993fk}, was derived for each site using $TWI= ln(\frac{a\alpha}{tan\theta})$, where, $\alpha$ is the upslope catchment area per unit width orthogonal to the flow direction and $\theta$ is the slope (in degrees). Across the 40 sites, TWI varied twofold (5 – 10 units). Altitude ranged by 200 m (from 1660 to 1860 m a.s.l), or by approximately 1.5\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}. This altitudinal temperature range is comparable to that observed in the OTC experiment (1.5\textdegree C vs 1.2\textdegree C, respectively).

@@

\section*{Data analysis.}
In total we built 11 hierarchical models to examine factors influencing tussock grass post-fire recovery and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models with Markov chain Monte Carlo (MCMC) sampling in R 3.2.2 using package Rstan 2.8.1.  For each model, three independent chains were executed. In all cases modelled parameters converged within 1000 iterations. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. After discarding the first 1000 iterations as ‘warm up’, a further 1000 iterations were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. The residuals of all models were checked graphically against predictions with none showing any systematic pattern or severe heteroscedasticity. To ensure all covariate effects were identifiable we selectively removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine-Shrub-Experiment}.

\subsection{\textit{Modelling shrub seedling growth rates.}}
For the purpose of modelling seedling annual growth rates we used heights measured in May of each year from seedlings that survived throughout the passive warming experiment. This meant that our observations were of maximum height for each individual $i$ at each May census $t$. We modelled these heights, $height_{it}$, as a random realization from a normal distribution truncated at zero:

\begin{equation}\label{eq:growth_obs} 
height_{it} \sim \mathcal{N^+}(\widehat{height}_{it}, \sigma_{obs}),
\end{equation}

where $\widehat{height}_{i,t}$ is the estimated maximum height for each individual $i$ at each May census $t$ and $\sigma_obs$ is the standard deviation of the observation error. The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:

@@

We used a logistic equation because it can simulate sigmoidal growth. That is, a curve whereby initial seedling growth is slow before increasing in speed and finally levelling off when adult height is reached. We used this particular parameterisation because it accounts for seedling size at the beginning of an experiment and contained biologically interpretable parameters. Here, $Hmax_{j[i]}$ is the asymptotic height: the average mature height an individual $i$ belonging to species $j$ can achieves. We treated Hmax as a constant, not to be estimated by the model, owing to our experimental observations being restricted to the first five years of growth. Based on maximum height data from 5 to 22 long-unburnt sites (depending on the species) we estimated the average mature height (and thus Hmax) to be approximately 34, 67, 66 and 70 cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively. $Hinit_i$ is the height of individual $i$ at the start of the experiment (i.e. the first May measurement). $Days_t$ is the number of days from the beginning of the experiment to each May census $t$.  The experiment ran for a total of 1812 days (or approximately 5 years). We divide the number of days at each census by 365.25 in order to estimate annual growth rates. The parameter of interest, $R_i$, is the mean growth rate constant for each individual $i$. This parameter controls the slope at inflection (i.e. the maximum annual relative growth rate). When multiplied by the asymptotic height, $Hmax$, it yields the maximum annual growth rate at the inflection point . We modelled $R_i$ separately for seedlings growing in bare ground (eqn. 1.3.1) compared to those growing in inter-tussock gaps (eqn. 1.3.2).

@@

The structure and parameter definitions differ slightly to that in eqn 1.3.1. Here, the model has two additional parameters, $\beta_2$ and $\beta_3$. The parameter $\beta_2$ is the effect of inter-tussock gap radius, which varied over time $t$, while $\beta_3$ is the interactive effect of OTC and time-varying inter-tussock gap radius. The inclusion of these parameters, coupled $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (12 cm). The parameter  $\beta_1$ is the OTC effect and $\epsilon_{plot_k}$ and $\epsilon_i$ are residual errors at the plot and individual level, respectively. Note that growth constant, $R$, now varies through time as a function of inter-tussock gap radius.

@@

We modelled individual instantaneous hazard rate, $\lambda_i$, in the same way we modelled $R_i$ in equations 1.3.1 \& 1.3.2. 
For individual seedlings of the four shrub species located in the bare ground patch of each plot, we modelled $\lambda_i$ using a varying intercept model:

@@

Here, the structure and parameter definitions are similar to eqn 2.3.1. The intercept, $\alpha$, represents the expected instantaneous hazard rate in control plots in an average inter-tussock gap radius, $\beta_1$ is the effect of OTC, $\beta_2$ is the effect of inter-tussock gap radius, $\beta_3$ is the interactive effect of OTC and inter-tussock gap radius and $\epsilon_{plot_k}$ and $\epsilon_i$ are the residual variation at the plot and individual scales.

@@

where $\mu_{k,t}$ is the log expected recruit count for each plot, and $\phi$ is the overdispersion parameter. We used a negative binomial distribution because count data was significantly overdispersed with $\phi$ being estimated as 3.95.

@@

\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
In each model, OTC treatment, $OTC_k$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census t. However, because some cardinal distances were recorded as censored data (i.e.  > 25 cm), these data were incomplete. We estimated these censored data using a censored model (see below). Once censored data were estimated, we then estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each censuses, for each inter-tussock seedling. Median inter-tussock gap radius ranged from 0.5 cm to 43 cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero (minus grand mean) and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as average responses and slope terms as partial dependencies conditional on other continuous variables at their mean. Furthermore, standardizing by two standard deviations allows the magnitude of effects to be directly compared between binary and continuous covariates\cite{Gelman:2007te}.

\subsection{\textit{Modelling censored Poa distances.}}
Approximately 15\% of the Poa cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, ≥ 26 cm.  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects and then using these distributions to estimate censored observations.

@@

where $\mu_i$ is the expected log estimated distance for cardinal distance $i$, and  $\sigma_obs$ is the standard deviation of the observation error.

@@

Again, like other models, half Cauchy priors with a scale of 25 were used for all standard deviation parameters. Both the intercept, $\alpha$, and the effect of OTC, $\beta$ (which are on the log scale), were estimated from weakly informed normal distributions centered on zero with a standard deviation of 2.5. Specifically, this prior encapsulated our knowledge that distances were very unlikely to be > 150 cm.

@@

\begin{equation}\label{eq:censored_obs} 
censored\;distance_i \sim LN(\mu_i, \sigma_{obs})[26,100],
\end{equation}

@@

\subsection{\textit{Modelling rates of inter-tussock gap infilling by Poa.}}
As well as modelling how OTC affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps. To examine this we first modelled the median gap radius,$y_{i,t}$, for each inter-tussock gap $i$ in each census $t$, as a random realisation from a lognormal distribution:

@@

where $\mu_i$ is the log expected seedling count for each plot, and $\phi$ is the overdispersion parameter. We used a negative binomial distribution because count data was significantly overdispersed for both species with $\phi$ being estimated as 0.54 ($\pm$ 0.1 95\% BCI) and 0.82 ($\pm$ 0.1 95\% BCI), for \textit{Grevillea} and \textit{Asterolasia}, respectively.

>>>>

\section*{Study site}
Research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals\cite{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003\cite{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests\cite{Williams:2008ba}.

\section*{Open Top Chamber Experiment}
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of grass recovery and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affect shrub seedling demographic rates, and 3) whether such affects change under warmer conditions post-fire conditions.

In alpine and tundra regions, shrub seedlings typically establish and grow in bare ground created by disturbance\cite{Williams:1987ua, Batllori:2009fe, Frost:2013tm}. Here, we used fire to create bare ground because 1) Australian alpine shrub seedlings occur in greatest abundance after fire (see seedling survey results) and 2) in many parts of the world, including Australia, wildfires are predicted to increase in frequency and severity\cite{Westerling:2006ve,Flannigan:2009kl}. Thus, to forecast likely changes in shrub abundance we need to know how warming affects growth and mortality rates of shrub seedlings growing in burnt sites.

We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient seedling numbers from species occurring in both open and closed heathland. We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all: obligate seeders post-fire (i.e. fire killed), common in the Australian Alps, and under climatic warming, likely candidates to invade non-shrubby plant communities such as alpine grasslands and herbfields. For each species, we collected a range of seedling heights (\Sexpr{range_init_size(complete_otc_seedling_data)[1]} to \Sexpr{range_init_size(complete_otc_seedling_data)[2]} cm) using a forestry planter (12.5 cm deep by 5.5 cm wide). The size of the planter was selected to minimize root damage across the range of seedling sizes and to allow for the extraction of a constant volume (378 cm$^3$) of intact soil and roots. The planter also minimized soil disturbance during seedling extraction and replanting.

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea} seedlings were transplanted into various sized inter-tussock gaps surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. S8). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

To simulate near-term warmer conditions predicted under IPCC\cite{IPCC:2013vv}, we used Open Top Chambers (OTCs)\cite{Molau:1996vc}. OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm. The chambers were constructed following the International Tundra Experiment (ITEX) protocols\cite{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (\textless 5\%; Solar Components Corporation, USA).

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) randomly installed into four control plots and four OTC plots. Each Micro Station measured ambient temperature, relative humidity (both at 10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Fig. S5-7). OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C, and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\section*{Seedling survey}
We measured seedling density and maximum heights in open heathland across the Bogong High Plains. The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (a measure of plant available water), adult density, fire, and, if burnt, fire severity, influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland. A secondary aim was to validate growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.

The study used 40 open heathland sites\cite{Camac:2013eb} established immediately after the 2003 fires\cite{Williams:2006tl}. All sites have not been grazed by cattle since the 2003 fires. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years\cite{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitudinal from 1660 m to 1860 m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2011-12, 9-years post-fire, seedling density in each site was estimated using forty 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots\texttimes4 transects) per site, subsampling an area of 2000 m$^2$.

Different variables were recorded at the three sampling levels (site, transect and plot). For each site, altitude and TWI were derived from a 30 m resolution digital elevation model of the Bogong High Plains. TWI, which approximates the relative plant-available water based on a site’s position in the landscape\cite{Moore:1993fk}, was derived for each site using $TWI= ln(\frac{a\alpha}{tan\theta})$, where, $\alpha$ is the upslope catchment area per unit width orthogonal to the flow direction and $\theta$ is the slope (in degrees). Across the 40 sites, TWI varied twofold (5 to 10 units; higher units = wetter). Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m (i.e. from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l), or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}. This altitudinal temperature range is comparable to that observed in the OTC experiment.

@@

\section*{Data analysis.}
In total we built 11 hierarchical models to examine factors influencing tussock grass post-fire recovery and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models with sampling in R \Sexpr{paste(version$major, version$minor, sep='.')} using package Rstan \Sexpr{packageVersion("rstan")}. For each model, three independent chains were executed. In all cases modelled parameters converged within 1000 iterations. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. After discarding the first 1000 iterations as ‘warm up’, a further 1000 iterations were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. The residuals of all models were checked graphically against predictions with none showing any systematic pattern or severe heteroscedasticity. To ensure all covariate effects were identifiable we selectively removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

\subsection{\textit{Modelling shrub seedling growth rates.}}
For the purpose of modelling seedling annual growth rates we used heights measured in May of each year from seedlings that survived throughout the passive warming experiment. This meant that our observations were of maximum height for each individual $i$ at each May census $t$. We modelled these heights, $height_{i,t}$, as a random realization from a normal distribution truncated at zero:

\begin{equation}\label{eq:growth_obs} 
height_{i,t} \sim \mathcal{N}(\widehat{height}_{i,t}, \sigma_{obs}), 0 \leq height_{i,t} \leq \infty,
\end{equation}

where $\widehat{height}_{i,t}$ is the estimated maximum height for each individual $i$ at each May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error. The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:

@@

We used a logistic equation because it can simulate sigmoidal growth. That is, a curve whereby initial seedling growth is slow before increasing in speed and finally levelling off when adult height is reached. We used this particular parameterisation because it accounts for seedling size at the beginning of an experiment and contained biologically interpretable parameters. Here, $Hmax_{j[i]}$ is the asymptotic height: the average mature height an individual $i$ belonging to species $j$ can achieves. We treated Hmax as a constant, not to be estimated by the model, owing to our experimental observations being restricted to the first five years of growth. Based on maximum height data from 5 to 22 long-unburnt sites (depending on the species) we estimated the average mature height (and thus Hmax) to be approximately \Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} and \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively. $Hinit_i$ is the height of individual $i$ at the start of the experiment (i.e. the first May measurement). $Days_t$ is the number of days from the beginning of the experiment to each May census $t$.  The experiment ran for a total of \Sexpr{n_days(chamber_dates, FALSE)} days (or approximately \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years). We divide the number of days at each census by 365.25 in order to estimate annual growth rates. The parameter of interest, $R_i$, is the mean growth rate constant for each individual $i$. This parameter controls the slope at inflection (i.e. the maximum annual relative growth rate). When multiplied by the asymptotic height, $Hmax$, it yields the maximum annual growth rate at the inflection point. We modelled $R_i$ separately for seedlings growing in bare ground (equation \ref{eq:r_bg}) compared to those growing in inter-tussock gaps (equation \ref{eq:r_tussock}).

@@

The structure and parameter definitions differ slightly to that in equation \ref{eq:r_bg}. Here, the model has two additional parameters, $\beta_2$ and $\beta_3$. The parameter $\beta_2$ is the effect of inter-tussock gap radius, which varied over time $t$, while $\beta_3$ is the interactive effect of OTC and time-varying inter-tussock gap radius. The inclusion of these parameters, coupled $Gap_{i,t}$ being centered and standardized (see Data inputs into growth and mortality models), alters the definition of the intercept, $\alpha$, which now represents the mean control plot growth rate at the average inter-tussock gap size (\Sexpr{mean_init_gap(tussock_otc_growth_data)} cm). The parameter  $\beta_1$ is the OTC effect and $\epsilon_{plot_k}$ and $\epsilon_i$ are residual errors at the plot and individual level, respectively. Note that growth constant, $R$, now varies through time as a function of inter-tussock gap radius.

@@

We modelled individual instantaneous hazard rate, $\lambda_i$, in the same way we modelled $R_i$ in equations \ref{eq:r_bg} \& \ref{eq:r_tussock}. 
For individual seedlings of the four shrub species located in the bare ground patch of each plot, we modelled $\lambda_i$ using a varying intercept model:

@@

Here, the structure and parameter definitions are similar to equation \ref{eq:r_tussock}. The intercept, $\alpha$, represents the expected instantaneous hazard rate in control plots in an average inter-tussock gap radius, $\beta_1$ is the effect of OTC, $\beta_2$ is the effect of inter-tussock gap radius, $\beta_3$ is the interactive effect of OTC and inter-tussock gap radius and $\epsilon_{plot_k}$ and $\epsilon_i$ are the residual variation at the plot and individual scales.

@@

where $\mu_{k,t}$ is the log expected recruit count for each plot, and $\phi$ is the overdispersion parameter. We used a negative binomial distribution because count data was significantly overdispersed with $\phi$ being estimated as \Sexpr{phi_value(recruit_model,'mean')} (95\% BCI: \Sexpr{phi_value(recruit_model,'l_bci')}\textendash \Sexpr{phi_value(recruit_model,'u_bci')}).

@@

\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
In each model, OTC treatment, $OTC_k$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census t. However, because some cardinal distances were recorded as censored data (i.e. \textgreater 25 cm), these data were incomplete. We estimated these censored data using a censored model (see below). Once censored data were estimated, we then estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each censuses, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as average responses and slope terms as partial dependencies conditional on other continuous variables at their mean. Furthermore, standardizing by two standard deviations allows the magnitude of effects to be directly compared between binary and continuous covariates\cite{Gelman:2007te}.

\subsection{\textit{Modelling censored Poa distances.}}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the Poa cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, ≥ 26 cm.  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects and then using these distributions to estimate censored observations.

@@

where $\mu_i$ is the expected log estimated distance for cardinal distance $i$, and  $\sigma_{obs}$ is the standard deviation of the observation error.

@@

Again, like other models, half Cauchy priors with a scale of 25 were used for all standard deviation parameters. Both the intercept, $\alpha$, and the effect of OTC, $\beta$ (which are on the log scale), were estimated from weakly informed normal distributions centered on zero with a standard deviation of 2.5. Specifically, this prior encapsulated our knowledge that distances were very unlikely to be \textgreater 150 cm.

@@

\begin{equation}\label{eq:censored_obs} 
censored\;distance_i \sim LN(\mu_i, \sigma_{obs}), 26 \leq censored\;distance_i \leq 100.
\end{equation}

@@

\subsection{\textit{Modelling rates of inter-tussock gap infilling by Poa.}}
As well as modelling how OTC affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps. To examine this we first modelled the median gap radius, $y_{i,t}$, for each inter-tussock gap $i$ in each census $t$, as a random realisation from a lognormal distribution:

@@

where $\mu_i$ is the log expected seedling count for each plot, and $\phi$ is the overdispersion parameter. We used a negative binomial distribution because count data was significantly overdispersed for both species with $\phi$ being estimated as \Sexpr{phi_value(greaus_seedling_density_model$fit,'mean')} (95\% BCI: \Sexpr{phi_value(greaus_seedling_density_model$fit,'l_bci')}\textendash \Sexpr{phi_value(greaus_seedling_density_model$fit,'u_bci')}) and \Sexpr{phi_value(asttry_seedling_density_model$fit,'mean')} (95\% BCI: \Sexpr{phi_value(asttry_seedling_density_model$fit,'l_bci')}\textendash \Sexpr{phi_value(asttry_seedling_density_model$fit,'u_bci')}), for \textit{Grevillea} and \textit{Asterolasia}, respectively.
